---
title: "Homework 6"
author: "Allan Kimaina"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)


# load GLM packages
library(pROC)
library(arm)
library(foreign)
library(nnet)
library(VGAM)
library(ordinal)
library(ModelGood)
library(InformationValue)
library(rms)
library(AER) 
library(metRology)
library(hett)
library(lme4)





riskyBehaviours.df=read.csv("data/risky_behaviors.csv")
riskyBehaviours.df$couples <- factor(riskyBehaviours.df$couples)
riskyBehaviours.df$women_alone <- factor(riskyBehaviours.df$women_alone)
riskyBehaviours.df$cBupacts <- (riskyBehaviours.df$bupacts - mean(riskyBehaviours.df$bupacts)) / (2 * sd(riskyBehaviours.df$bupacts))

# cdf
# load required data
cd4.df=read.csv("data/cd4.csv")
cd4.df$newpid= as.character(cd4.df$newpid)
cd4.df$treatmnt= as.factor(cd4.df$treatmnt)

# transfrom data
time<-cd4.df$visage-cd4.df$baseage
cd4.df<-cbind(cd4.df,time)


```

\onecolumn

# Chapter 13 Question 5:
**Return to the CD4 data introduced from Exercise 11.4.**

## Part A:
``Extend the model in Exercise 12.2 to allow for varying slopes for the time predictor.``

```{r results='asis', echo=FALSE, warning=FALSE}
mod13q5a = lmer(cd4pct~1+time+treatmnt+baseage+(1+time|newpid),data=cd4.df,REML=F)

summary(mod13q5a)
#coef(model_chap13_q5a)


```


## Part B:
`` Next fit a model that does not allow for varying slopes but does allow for different coefficients for each time point (rather than fitting the linear trend)``

```{r results='asis', echo=FALSE, warning=FALSE}

cd4.df = cd4.df %>%
          mutate(
            time_point = case_when(
              time == 0 ~ 0,
              time > 0 & time <= .25+0.125  ~ 1,
              time > .25+0.125 & time <= .5+ 0.125 ~ 2,
              time > .5+ 0.125 & time <= .75+ 0.125 ~ 3,
              time > .75+ 0.125 & time <= 1+ 0.125 ~ 4, 
              time > 1+ 0.125 & time <= 1.25+ 0.125 ~ 5,
              time > 1.25+ 0.125 & time <= 1.5+ 0.125 ~ 6, 
              time > 1.5+ 0.125 & time <= 1.75+ 0.125 ~ 7,
              time > 1.75+ 0.125  ~ 8
            )
          )


mod13q5b = lmer(cd4pct~1+treatmnt+baseage+(1|time_point),data=cd4.df,REML=F)

summary(mod13q5b)
```



## Part C:
`` Compare the results of these models both numerically and graphically.``


### Graphically

```{r results='asis', echo=FALSE, warning=FALSE}

```




### Numerically
```{r results='asis', echo=FALSE, warning=FALSE}

quest13b <- anova(mod13q5b,mod13q5a)

```





\onecolumn

# Chapter 14 Question 3:
**Three-level logistic regression: the folder rodents contains data on rodents in a sample of New York City apartments.**
```{r results='asis', echo=FALSE, warning=FALSE}

df <- read.table("data/rodents.txt")
df$race <- factor(df$race, labels=c("White (non-hispanic)", "Black (non-hispanic)", "Puerto Rican", "Other Hispanic", "Asian/Pacific Islander", "Amer-Indian/Native Alaskan", "Two or more races"))

df$unitflr2 <- as.factor(df$unitflr2)
df$numunits <- as.factor(df$numunits)
df$stories <- as.factor(df$stories)
df$extwin4_2 <- as.factor(df$extwin4_2)
df$extflr5_2 <- as.factor(df$extflr5_2)
df$borough <- factor(df$borough, labels=c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"))
df$cd <- as.factor(df$cd)
df$intcrack2 <- as.factor(df$intcrack2)
df$inthole2 <- as.factor(df$inthole2)
df$intleak2 <- as.factor(df$intleak2)
df$intpeel_cat <- as.factor(df$intpeel_cat)
df$help <- as.factor(df$help)
df$old <- as.factor(df$old)
df$dilap <- as.factor(df$dilap)
df$regext <- as.factor(df$regext)
df$poverty <- as.factor(df$poverty)
df$povertyx2 <- as.factor(df$povertyx2)
df$housing <- factor(df$housing, labels=c("public", "rent controlled/stabilized", "owned", "other rentals"))
df$board2 <- as.factor(df$board2)
df$subsidy <- as.factor(df$subsidy)
df$under6 <- as.factor(df$under6)


# Missing values
missingNA <- sapply(df, function(x) sum(is.na(x)))
df <- na.omit(df)


```

## Part A:
``Build a varying intercept logistic regression model (varying over buildings) to predict the presence of rodents (the variable rodent2 in the dataset) given indicators for the ethnic groups (race) as well as other potentially relevant predictors describing the apartment and building. Fit this model using lmer() and interpret the coefficients at both levels. ``

```{r results='asis', echo=FALSE, warning=T}

model.14.3A <- glmer(rodent2 ~ 1+race+personrm +intcrack2 + inthole2 + intleak2 +
                            struct +regext+extflr5_2 +
                           # old+dilap+intpeel_cat+extwin4_2+housing +
                            (1|bldg),
                            data=df,
                            family=binomial(link="logit"),
                            control=glmerControl(
                                                  optimizer="bobyqa",
                                                 optCtrl=list(maxfun=200000)) 
                            )
summary(model.14.3A)

```


## Part B:
``Now extend the model in (b) to allow variation across buildings within community district and then across community districts. Also include predictors describing the community districts. Fit this model using lmer() and interpret the coefficients at all levels.``

```{r results='asis', echo=FALSE, warning=T}
model.14.3B <- glmer(rodent2 ~ 1+race+personrm +intcrack2 + inthole2 + intleak2 +
                            struct +regext+extflr5_2 +
                           # old+dilap+intpeel_cat+extwin4_2+housing +
                            (1|bldg)+
                            (1|cd),
                            data=df,
                            family=binomial(link="logit"),
                            # increase convergence iterations
                            control=glmerControl(
                                                  optimizer="bobyqa",
                                                 optCtrl=list(maxfun=200000)) 
                            )
summary(model.14.3B)

```

## Part C:
`` Compare the fit of the models in (a) and (b).``

```{r results='asis', echo=FALSE, warning=T}

anova_logit.14 <- anova(model.14.3B,model.14.3A)

```



\onecolumn

# Source Code

```{r echo=T, warning=F}


```
